---
title: "Parallel Computing in R"
output: html_document
runtime: shiny_prerendered
---
  
```{r setup, include=FALSE}
library(learnr)
library(parallel)
knitr::opts_chunk$set(echo = FALSE)
```

## **Introduction to Parallelization**

Parallel computing in R allows us to perform computations in parallel, utilizing multiple cores or distributed computing resources. The "parallel" package provides functionality for parallelization in R.

To begin, let's explore a simple example of parallel computation using the "parallel" package. The parLapply() function is useful for applying a function in parallel across a list.

In the code chunk below, we have a list of numbers and a function that doubles each number. Use parLapply() to apply the function to the list in parallel.

```{r parallel_1a, exercise = TRUE}

numbers <- list(1, 2, 3, 4, 5)
double <- function(x) {
  return(x * 2)
}

# Create a cluster object with the desired number of workers
cl <- makeCluster(2)

# Perform parallel computation using parLapply on the cluster
result <- parLapply(cl, numbers, double)

# Print the result
print(result)

# Stop the cluster
stopCluster(cl)

```

**Independent Exercise**

In the code chunk below, create a list named my_list containing the numbers 6, 7, 8, 9, and 10. Write a function named square that squares each number in the list. Use parLapply() to apply the function to the list in parallel. Print the result.

```{r parallel_1b, exercise = TRUE, exercise.lines = 8}
```

## **Parallelization Strategies**

The "parallel" package provides various functions for parallelization, allowing us to choose the appropriate strategy based on our requirements.

### **Fork-Based Parallelization**

Fork-based parallelization is suitable for multicore machines. It utilizes the available cores to perform computations in parallel. The mclapply() function is a parallelized version of the lapply() function.

In the code chunk below, we have a list of numbers and a function that calculates the cube of each number. Use mclapply() to apply the function to the list in parallel.

```{r parallel_1c, exercise = TRUE}

numbers <- list(1, 2, 3, 4, 5)
cube <- function(x) {
  x^3
}

# Create a cluster object with the desired number of workers
cl <- makeCluster(detectCores())

# Perform parallel computation using mclapply on the cluster
result <- mclapply(numbers, cube)

# Print the result
print(result)

# Stop the cluster
stopCluster(cl)

```

**Independant Exercise**

In the code chunk below, create a list named my_list containing the numbers 36, 49, 64, 81, and 100. Write a function named square_root that calculates the square root of each number in the list. Use mclapply() to apply the function to the list in parallel. Print the result.

```{r parallel_1d, exercise = TRUE, exercise.lines = 8}
```

### **Socket-Based Parallelization**

Socket-based parallelization allows us to perform computations across multiple machines in a network. The clusterApply() function is useful for applying a function in parallel across a cluster.

In the code chunk below, we have a list of numbers and a function that calculates the factorial of each number. Create a cluster of two workers using the makeCluster() function. Use clusterApply() to apply the function to the list in parallel.

```{r parallel_1e, exercise = TRUE}

numbers <- list(5, 6, 7, 8, 9)
factorial <- function(x) {
  return(prod(1:x))
}

cl <- makeCluster(2)
result <- clusterApply(cl, numbers, factorial)
print(result)

stopCluster(cl)

```
**Independent Exercise**

In the code chunk below, create a list named my_list containing the numbers 3, 4, 5, 6, and 7. Write a function named power_of_two that calculates 2 raised to the power of each number in the list. Create a cluster of three workers using the makeCluster() function. Use clusterApply() to apply the function to the list in parallel. Print the result.

```{r parallel_1f, exercise = TRUE, exercise.lines = 8}
```
